Windowed residual-minimization model reduction for dynamical systems

### 2019-10-02 main.tex
p 6

The layout of the rest of this paper is as follows: Section~\ref{sec:math} outlines the problem statement of interest, along with the Galerkin, LSPG, and space--time LSPG ROMs. Section~\ref{sec:tclspg} outlines the \methodAcronym\ formulation proposed in this work. Section~\ref{sec:numerical_techniques} outlines numerical techniques for solving \methodAcronymROMs. Both direct and indirect methods are discussed. Next, Section~\ref{sec:analysis} provides equivalency conditions and error analysis for \methodAcronymROMs. Section~\ref{sec:numerical_experiments} presents numerical experiments. Conclusions and perspectives are provided in Section~\ref{sec:conclude}.
The notation used is as follows: We denote vector functionals in stylized bold text, i.e., $\boldsymbol x$. We denote vectors in standard bold text, i.e., $\mathbf{x}$. Matrices are denoted with capital bold letters, i.e., $\mathbf{X} \equiv \begin{bmatrix} \mathbf{x}_1 & \cdots & \mathbf{x}_r\end{bmatrix}$. Vector and functional spaces are denoted with calligraphic fonts, i.e., $\mathcal{X}$.
%Model reduction of nonlinear dynamical systems is a growing research area.
%The least-squares Petrov-Galerkin (LSPG) approach has gained popularity as a model reduction technique for nonlinear dynamical systems. Derived at the fully discrete level, the LSPG approach solves a least-squares problem that minimizes the fully discrete residual of the reduced-order model at each time-step. LSPG has been shown to be an effective formulation for nonlinear model reduction on complex problems of interest.
%\begin{comment} 
\section{Mathematical Setting}\label{sec:math}
We consider the system of nonlinear ordinary differential equations (ODEs),
\begin{equation}\label{eq:FOM}
 \frac{d}{dt}\stateFOMArg{}{t} = \velocity(\stateFOMArg{}{t},t), \qquad \stateFOM(0) = \stateFOMIC, \qquad t \in [0,T],
\end{equation}
where $T \in \mathbb{R}^+$ is the final time, $\stateFOM \in \RR{\fomdim} \otimes \timeSpace $ with $\stateFOM: [0,T] \rightarrow  \RR{\fomdim}$ denotes the state implicitly defined to be the solution to~\eqref{eq:FOM}, $\timeSpace$ is the set of functions acting on $[0,T]$, $\stateFOMIC \in \mathbb{R}^{\fomdim}$ are the initial conditions, and $\velocity: \mathbb{R}^{\fomdim} \times [0,T] \rightarrow \mathbb{R}^{\fomdim}$ with $(\stateyDiscrete,t) \mapsto \velocity(\stateyDiscrete,t)$ is the velocity. Equation~\eqref{eq:FOM} will be referred to as the ``full-order model" (FOM) ODE.

Directly solving~\eqref{eq:FOM} can be computationally expensive due to both the high dimensionality of the state and the range of time-scales that needs to be resolved. For time-critical applications, it is important to devise computationally efficient approximations to~\eqref{eq:FOM}. Projection-based ROMs are 
one such promising approach.

\subsection{Projection-Based Reduced-Order Models}
Projection-based ROMs generate approximate solutions to the FOM ODE~\eqref{eq:FOM} by approximating the state in a low-dimensional trial subspace. For dynamical 
systems, two distinct types of trial subspaces are used: space--time trial subspaces with spatial dimension reduction (\spatialAcronym) and space--time trial subspaces with space--time dimension reduction (\spaceTimeAcronym). \spatialAcronym\ trial subspaces are most commonly employed and are used in, e.g., the Galerkin and Least-Squares Petrov--Galerkin (LSPG) methods. \spaceTimeAcronym\ trial subspaces are typically 
employed in space--time ROMs. The following subsections outline these two trial subspaces, along with the Galerkin, LSPG, and 
space--time LSPG method.

\subsection{\spatialAcronym\ Trial Subspaces}
\spatialAcronym\ trial subspaces approximate the FOM ODE solution, $\stateFOM$, with an approximate solution, $\approxstate$, that lives in a $K$-dimensional affine trial subspace, 
\begin{align}\label{eq:spatial_subspace}
& \stspaceS \defeq \trialspace \otimes \timeSpace,\\ 
& \trialspace \defeq \text{Range}(\basismat) +  \stateIntercept,  \label{eq:spatial_subspace2}  \\
&\basismat \defeq \begin{bmatrix}  \basisvec_1  & \cdots &  \basisvec_K \end{bmatrix}, \label{eq:spatial_subspace3} 
\end{align}
where we refer to $\basismat \in \RRStar{\romdim}{\fomdim}$, with $\RRStar{\romdim}{\fomdim}$ denoting the set  $\{ \mathbf{X} \in \RR{\fomdim \times \romdim} \big| \mathbf{X}^T \mathbf{X} = \mathbf{I} \}$, as the spatial trial basis matrix, $\trialspace$ as the spatial trial subspace, and $\stateIntercept \in \mathbb{R}^{\fomdim}$ as the reference state. 
%where $\basisvec_i \in \mathbb{R}^N$ are orthonormal basis vectors and $\stateIntercept \in \mathbb{R}^{\fomdim}$ the ``intercept''.
%We refer to $\basismat \in \RRStar{\fomdim \times \romdim}$, with $\RRStar{p \times q}$ denoting the set of orthogonal $q$ frames in $p$, as the spatial trial basis, and $\trialspace$ as the spatial trial space. 
The basis vectors $\basisvec_i$ are typically constructed through, e.g., POD~\cite{berkooz_turbulence_pod}, the reduced-basis method~\cite{rb_1,rb_2,rb_3,NgocCuong2005,Rozza2008}, etc. 
ROMs leveraging \spatialAcronym\ trial subspaces approximate the state as,
\begin{equation}\label{eq:affine_trialspace}
\stateFOMArg{}{t} \approx \approxstate(t) \defeq \basisspace \genstateArg{}{t} + \stateIntercept,
\end{equation}
where $\genstate \in \RR{\romdim} \otimes \timeSpace$, with $\genstate : [0,T] \rightarrow \mathbb{R}^{\romdim}$, are the generalized coordinates. Substituting the approximation given in~\eqref{eq:affine_trialspace} into~\eqref{eq:FOM} and performing orthogonal $\elltwo$ projection of the initial conditions onto the trial subspace, one obtains,
\begin{equation}\label{eq:g_truncation}
\basisspace \frac{d }{dt}\genstateArg{}{t} = \velocity(\basisspace
\genstateArg{}{t} + \stateIntercept,t ), \qquad \genstate(0) = \genstateIC, \qquad t \in [0,T].
\end{equation}
Equation~\eqref{eq:g_truncation} provides an over-determined system of $N$ equations for the $K$ generalized coordinates. Dimension reduction is predominantly performed with either the \textit{Galerkin method} or the \textit{Least-Squares Petrov--Galerkin method} (LSPG). We now outline these two methods.

\subsubsection{The Galerkin Method}

@@

\subsubsection{Least-Squares Petrov--Galerkin}

@@

\subsection{\spaceTimeAcronym\ Trial Spaces and Space--Time ROMs}
Recent work has investigated space--time reduced-order models~\cite{choi_stlspg,constantine_strom,URBAN2012203,Yano2014ASC,benner_st,bui_thesis}. Space--time methods 
are advantageous in that they allow for both spatial \textit{and} temporal dimension reduction. Further, space--time 
methods are often equipped with more favorable error bounds than non-space--time methods. The approach proposed in this work displays commonalities 
with space--time methods and, as such, we briefly outline \spaceTimeAcronym\ trial subspaces and space--time LSPG.

@@

Substituting the approximation~\eqref{eq:stapprox1} into the FOM ODE~\eqref{eq:FOM},
\begin{equation}\label{eq:st_fom}
\frac{d }{dt} \stbasisArg{}{t} \stgenstate =  \velocity(\stbasisArg{}{t} \stgenstate + \stateIntercept,t) , \qquad \genstate(0) =\mathbb{P}(\stateFOM)(0), \qquad t \in [0,T],
\end{equation}
where $\mathbb{P} : \RR{N} \otimes \timeSpace \rightarrow \stspaceST$ is an orthogonal projection onto the trial subspace (e.g., $\elltwo$).
Equation~\eqref{eq:st_fom} is an over-determined algebraic system for the $\stdim$ unknowns. We now outline the Space--Time Least-Squares Petrov--Galerkin method~\cite{choi_stlspg}. 
The space--time Galerkin method is noted to additionally be a possibility for model reduction. As the space--time Galerkin method does not associate with any residual minimization 
principle, however, we do not outline it here.
\subsubsection{Space--Time LSPG} 
As in LSPG, Space--Time LSPG (ST-LSPG)~\cite{choi_stlspg} minimizes the fully discrete residual, with the key difference being that ST-LSPG minimizes the residual over the entire time domain. For brevity, ST-LSPG is outlined only for linear multistep methods.
%As ST-LSPG acts on the discrete solution, we first note that the discrete FOM solution lives in $\RR{N} \otimes \RR{N_t}$. ST-LSPG then seeks solutions in the trial space,
%$$\stspaceDiscrete \defeq \spaceTrialSpace \otimes \timeTrialSpaceDiscrete \subseteq \RR{N} \otimes \RR{N_t},$$
%where $\timeTrialSpaceDiscrete$ is the discrete temporal trial space. It is noted that this trial space is a specific instance of the trial space~\ref{eq:sttrialspace_def} 
%As before, we define a discretization of the continuous time-domain $[0,T]$ into $N_t$ time instances $t^n = n\Delta t, n=1,2,\ldots,N_t$ with a constant time-step $\Delta t = T/N_t$. We define now the space--time discrete residual for the implicit Euler method as,
We first define the space--time discrete residual for linear multistep methods,
\begin{align*}
\residST_{\text{LMS}} & \vcentcolon (\stateyDiscreteArg{N_t}, \ldots ,\stateyDiscreteArg{1};\stateFOMIC) \mapsto \begin{bmatrix}
%\frac{\stateyDiscreteArg{N_t} - \stateyDiscreteArg{N_t-1}  }{\Delta t} -  \velocity(\stateyDiscreteArg{N_t}) \\
\residLMSArg{N_t}(\stateyDiscreteArg{N_t};\stateyDiscreteArg{N_t-1},\ldots,\stateyDiscreteArg{N_t - k^{N_t}}) \\ 
%\frac{\stateyDiscreteArg{N_t-1} - \stateFOMDiscreteArg{N_t-2}  }{\Delta t} -  \velocity(\stateyDiscreteArg{N_t-1}) \\
\vdots \\
%\frac{\stateyDiscreteArg{1} - \stateFOMIC }{\Delta t} -  \velocity(\stateyDiscreteArg{1}) \\
\residLMSArg{1}(\stateyDiscreteArg{1}; \stateFOMIC) 
\end{bmatrix}  \\
& \vcentcolon \RR{\fomdim} \otimes \RR{N_t+1} \rightarrow \RR{\fomdim N_t}. 
\end{align*}
For convenience we define the same residual with the generalized coordinates as the inputs,
\begin{align*}
\genresidST_{\text{LMS}} & \vcentcolon (\stgenstatey ;\stateFOMIC) \mapsto \begin{bmatrix}
%\frac{\stateyDiscreteArg{N_t} - \stateyDiscreteArg{N_t-1}  }{\Delta t} -  \velocity(\stateyDiscreteArg{N_t}) \\
\residLMSArg{n}\bigg( \stbasisArg{}{t^{N_t}}\stgenstatey + \stateIntercept; \stbasisArg{}{t^{N_t-1}}\stgenstatey + \stateIntercept ,\ldots,  \stbasisArg{}{t^{N_t-k^{N_t}}}\stgenstatey + \stateIntercept \bigg) \\ 
\vdots \\
\residLMSArg{n}\bigg( \stbasisArg{}{t^1}\stgenstatey + \stateIntercept; \stateFOMIC \bigg) \\ 
\end{bmatrix}  \\
& \vcentcolon \RR{\stdim} \times \RR{N} \rightarrow \RR{\fomdim N_t}. 
\end{align*}
ST-LSPG is defined as,
\begin{equation*}
\stgenstate = \underset{ \stgenstatey }{\text{arg\,min}}|| \lspgWeightingSTArg{\stgenstatey}  \genresidST_{\text{LMS}}(\stgenstatey;\stateFOMIC) ||_2^2, 
\end{equation*}
where $\lspgWeightingSTArg{\cdot} \in \RR{n_{st} \times N N_t}$, with $\stdim \le n_{st} \le N N_t$, is a space--time weighting matrix.

@@

\subsection{Outstanding Challenges}
This work seeks to overcome the limitations of existing reduced-order modeling approaches. Specifically, 
we look to overcome LSPG's time-step and time-scheme dependence, the stability challenges of both Galerkin and LSPG, and the potential computational cost of ST-LSPG. To this end, we now outline the \methodName\ formulation. 
%\end{comment}
%\input{formulation_3}

>>>>

The paper proceeds as follows: Section~\ref{sec:math}
outlines the mathematical setting for the full-order model, along with the Galerkin, LSPG, and
space--time LSPG ROMs; Section~\ref{sec:tclspg} outlines the proposed \methodAcronym\
formulation; Section~\ref{sec:numerical_techniques}
outlines numerical techniques for solving \methodAcronymROMs, including both direct and
indirect methods; Section~\ref{sec:analysis} provides
equivalence conditions and error analysis for \methodAcronymROMs;
Section~\ref{sec:numerical_experiments} presents numerical experiments;
Section~\ref{sec:conclude} provides conclusions and perspectives.
We denote vector-valued functions with stylized bold text (i.e., $\boldsymbol
x$), vectors with standard bold text (i.e., $\mathbf{x}$), 
matrices with capital bold letters (i.e., $\mathbf{X} \equiv \begin{bmatrix}
\mathbf{x}_1 & \cdots & \mathbf{x}_r\end{bmatrix}$), and spaces with
calligraphic fonts (i.e., $\mathcal{X}$).
%Model reduction of nonlinear dynamical systems is a growing research area.
%The least-squares Petrov-Galerkin (LSPG) approach has gained popularity as a model reduction technique for nonlinear dynamical systems. Derived at the fully discrete level, the LSPG approach solves a least-squares problem that minimizes the fully discrete residual of the reduced-order model at each time-step. LSPG has been shown to be an effective formulation for nonlinear model reduction on complex problems of interest.
%\begin{comment} 
\section{Mathematical setting}\label{sec:math}
We consider the system of nonlinear ordinary differential equations (ODEs),
\begin{equation}\label{eq:FOM}
 \frac{d\stateFOM}{dt} = \velocity(\stateFOM,t), \qquad \stateFOM(0) = \stateFOMIC,
\end{equation}
where $t \in [0,T]$ denotes time with 
$T \in \mathbb{R}^+$ denoting the final time, $\stateFOM \in \RR{\fomdim}
	\otimes \timeSpace $ with $\stateFOM: [0,T] \rightarrow  \RR{\fomdim}$
	denotes the state implicitly defined as the solution to initial value
	problem~\eqref{eq:FOM}, $\stateFOMIC \in \mathbb{R}^{\fomdim}$ denotes the initial conditions,
	and $\velocity: \mathbb{R}^{\fomdim} \times [0,T] \rightarrow
	\mathbb{R}^{\fomdim}$ with $(\stateyDiscrete,t) \mapsto
	\velocity(\stateyDiscrete,t)$ denotes the velocity, which is possibly
	nonlinear in its first argument. We refer to Eq.~\eqref{eq:FOM} 
	as the ``full-order model" (FOM) ODE.

Directly solving initial value problem~\eqref{eq:FOM} is computationally
	expensive if either the state-space dimension $\fomdim$ is large, or if the
	time-interval length $T$ is large relative to the time step required to
	numerically integrate Eq.~\eqref{eq:FOM}. For time-critical or many-query
	applications, it is essential to employ an approximation in lieu of the FOM
	ODE~\eqref{eq:FOM}. Projection-based ROMs constitute one such promising
	approach.

\subsection{Projection-based reduced-order models}
Projection-based ROMs generate approximate solutions to the FOM
	ODE~\eqref{eq:FOM} by approximating the state in a low-dimensional trial
	subspace. For dynamical 
	systems, two distinct types of space--time trial subspaces are commonly used: 
\begin{enumerate} 
	\item \textit{\spatialAcronym\ trial subspaces}}. These 
	trial subspaces reduce only the spatial dimension of the FOM, and are commonly used
		in model-reduction methods based on spatial projection followed by time
		integration (e.g., Galerkin projection, LSPG projection)
	\item \textit{\spaceTimeAcronym\ trial subspaces}. These 
	trial subspaces reduce both the spatial and temporal dimensions of the FOM,
		and are commonly employed 
		employed in model-reduction methods based on space--time projection (e.g.,
		space--time LSPG projection).
\end{enumerate}
 We now describe these two types of trial subspaces and their application to 
	Galerkin, LSPG, and space--time LSPG projection. 
	\KTC{HERE}

\subsection{\spatialAcronym\ trial subspaces}
\spatialAcronym\ trial subspaces approximate the FOM ODE solution, $\stateFOM$, with an approximate solution, $\approxstate$, that lives in a $K$-dimensional affine trial subspace, 
\begin{align}\label{eq:spatial_subspace}
& \stspaceS \defeq \trialspace \otimes \timeSpace,\\ 
& \trialspace \defeq \text{Range}(\basismat) +  \stateIntercept,  \label{eq:spatial_subspace2}  \\
&\basismat \defeq \begin{bmatrix}  \basisvec_1  & \cdots &  \basisvec_K \end{bmatrix}, \label{eq:spatial_subspace3} 
\end{align}
where we refer to $\basismat \in \RRStar{\romdim}{\fomdim}$, with $\RRStar{\romdim}{\fomdim}$ denoting the set  $\{ \mathbf{X} \in \RR{\fomdim \times \romdim} \big| \mathbf{X}^T \mathbf{X} = \mathbf{I} \}$, as the spatial trial basis matrix, $\trialspace$ as the spatial trial subspace, and $\stateIntercept \in \mathbb{R}^{\fomdim}$ as the reference state. 
%where $\basisvec_i \in \mathbb{R}^N$ are orthonormal basis vectors and $\stateIntercept \in \mathbb{R}^{\fomdim}$ the ``intercept''.
%We refer to $\basismat \in \RRStar{\fomdim \times \romdim}$, with $\RRStar{p \times q}$ denoting the set of orthogonal $q$ frames in $p$, as the spatial trial basis, and $\trialspace$ as the spatial trial space. 
The basis vectors $\basisvec_i$ are typically constructed through, e.g., POD~\cite{berkooz_turbulence_pod}, the reduced-basis method~\cite{rb_1,rb_2,rb_3,NgocCuong2005,Rozza2008}, etc. 
ROMs leveraging \spatialAcronym\ trial subspaces approximate the state as,
\begin{equation}\label{eq:affine_trialspace}
\stateFOMArg{}{t} \approx \approxstate(t) \defeq \basisspace \genstateArg{}{t} + \stateIntercept,
\end{equation}
where $\genstate \in \RR{\romdim} \otimes \timeSpace$, with $\genstate : [0,T] \rightarrow \mathbb{R}^{\romdim}$, are the generalized coordinates. Substituting the approximation given in~\eqref{eq:affine_trialspace} into~\eqref{eq:FOM} and performing orthogonal $\elltwo$ projection of the initial conditions onto the trial subspace, one obtains,
\begin{equation}\label{eq:g_truncation}
\basisspace \frac{d }{dt}\genstateArg{}{t} = \velocity(\basisspace
\genstateArg{}{t} + \stateIntercept,t ), \qquad \genstate(0) = \genstateIC, \qquad t \in [0,T].
\end{equation}
Equation~\eqref{eq:g_truncation} provides an over-determined system of $N$ equations for the $K$ generalized coordinates. Dimension reduction is predominantly performed with either the \textit{Galerkin method} or the \textit{Least-Squares Petrov--Galerkin method} (LSPG). We now outline these two methods.

\subsubsection{Galerkin projection}

@@

\subsubsection{Least-Squares Petrov--Galerkin projection}

@@

\subsection{\spaceTimeAcronym\ trial spaces and space--time ROMs}
Recent work has investigated space--time reduced-order models~\cite{choi_stlspg,constantine_strom,URBAN2012203,Yano2014ASC,benner_st,bui_thesis}. Space--time methods 
are advantageous in that they allow for both spatial \textit{and} temporal dimension reduction. Further, space--time 
methods are often equipped with more favorable error bounds than non-space--time methods. The approach proposed in this work displays commonalities 
with space--time methods and, as such, we briefly outline \spaceTimeAcronym\ trial subspaces and space--time LSPG.

@@

Substituting the approximation~\eqref{eq:stapprox1} into the FOM ODE~\eqref{eq:FOM},
\begin{equation}\label{eq:st_fom}
\frac{d }{dt} \stbasisArg{}{t} \stgenstate =  \velocity(\stbasisArg{}{t} \stgenstate + \stateIntercept,t) , \qquad \genstate(0) =\mathbb{P}(\stateFOM)(0), \qquad t \in [0,T],
\end{equation}
where $\mathbb{P} : \RR{N} \otimes \timeSpace \rightarrow \stspaceST$ is an orthogonal projection onto the trial subspace (e.g., $\elltwo$).
Equation~\eqref{eq:st_fom} is an over-determined algebraic system for the $\stdim$ unknowns. We now outline the Space--Time Least-Squares Petrov--Galerkin method~\cite{choi_stlspg}. 
The space--time Galerkin method is noted to additionally be a possibility for model reduction. As the space--time Galerkin method does not associate with any residual minimization 
principle, however, we do not outline it here.
\subsubsection{Space--Time LSPG projection}  
As in LSPG, Space--Time LSPG (ST-LSPG)~\cite{choi_stlspg} minimizes the fully discrete residual, with the key difference being that ST-LSPG minimizes the residual over the entire time domain. For brevity, ST-LSPG is outlined only for linear multistep methods.
%As ST-LSPG acts on the discrete solution, we first note that the discrete FOM solution lives in $\RR{N} \otimes \RR{N_t}$. ST-LSPG then seeks solutions in the trial space,
%$$\stspaceDiscrete \defeq \spaceTrialSpace \otimes \timeTrialSpaceDiscrete \subseteq \RR{N} \otimes \RR{N_t},$$
%where $\timeTrialSpaceDiscrete$ is the discrete temporal trial space. It is noted that this trial space is a specific instance of the trial space~\ref{eq:sttrialspace_def} 
%As before, we define a discretization of the continuous time-domain $[0,T]$ into $N_t$ time instances $t^n = n\Delta t, n=1,2,\ldots,N_t$ with a constant time-step $\Delta t = T/N_t$. We define now the space--time discrete residual for the implicit Euler method as,
We first define the space--time discrete residual for linear multistep methods,
\begin{align*}
\residST_{\text{LMS}} & \vcentcolon (\stateyDiscreteArg{N_t}, \ldots ,\stateyDiscreteArg{1};\stateFOMIC) \mapsto \begin{bmatrix}
%\frac{\stateyDiscreteArg{N_t} - \stateyDiscreteArg{N_t-1}  }{\Delta t} -  \velocity(\stateyDiscreteArg{N_t}) \\
\residLMSArg{N_t}(\stateyDiscreteArg{N_t};\stateyDiscreteArg{N_t-1},\ldots,\stateyDiscreteArg{N_t - k^{N_t}}) \\ 
%\frac{\stateyDiscreteArg{N_t-1} - \stateFOMDiscreteArg{N_t-2}  }{\Delta t} -  \velocity(\stateyDiscreteArg{N_t-1}) \\
\vdots \\
%\frac{\stateyDiscreteArg{1} - \stateFOMIC }{\Delta t} -  \velocity(\stateyDiscreteArg{1}) \\
\residLMSArg{1}(\stateyDiscreteArg{1}; \stateFOMIC) 
\end{bmatrix}  \\
& \vcentcolon \RR{\fomdim} \otimes \RR{N_t+1} \rightarrow \RR{\fomdim N_t}. 
\end{align*}
For convenience we define the same residual with the generalized coordinates as the inputs,
\begin{align*}
\genresidST_{\text{LMS}} & \vcentcolon (\stgenstatey ;\stateFOMIC) \mapsto \begin{bmatrix}
%\frac{\stateyDiscreteArg{N_t} - \stateyDiscreteArg{N_t-1}  }{\Delta t} -  \velocity(\stateyDiscreteArg{N_t}) \\
\residLMSArg{n}\bigg( \stbasisArg{}{t^{N_t}}\stgenstatey + \stateIntercept; \stbasisArg{}{t^{N_t-1}}\stgenstatey + \stateIntercept ,\ldots,  \stbasisArg{}{t^{N_t-k^{N_t}}}\stgenstatey + \stateIntercept \bigg) \\ 
\vdots \\
\residLMSArg{n}\bigg( \stbasisArg{}{t^1}\stgenstatey + \stateIntercept; \stateFOMIC \bigg) \\ 
\end{bmatrix}  \\
& \vcentcolon \RR{\stdim} \times \RR{N} \rightarrow \RR{\fomdim N_t}. 
\end{align*}
ST-LSPG is defined as,
\begin{equation*}
\stgenstate = \underset{ \stgenstatey }{\text{arg\,min}}|| \lspgWeightingSTArg{\stgenstatey}  \genresidST_{\text{LMS}}(\stgenstatey;\stateFOMIC) ||_2^2, 
\end{equation*}
where $\lspgWeightingSTArg{\cdot} \in \RR{n_{st} \times N N_t}$, with $\stdim \le n_{st} \le N N_t$, is a space--time weighting matrix.

@@

\subsection{Outstanding challenges}
This work seeks to overcome the limitations of existing reduced-order modeling approaches. Specifically, 
we look to overcome LSPG's time-step and time-scheme dependence, the stability challenges of both Galerkin and LSPG, and the potential computational cost of ST-LSPG. To this end, we now outline the \methodName\ formulation. 
%\end{comment}
%\input{formulation_3}

### 2019-10-02 section4/numerical_methods.tex
p 6

\section{Numerical Solution Techniques}\label{sec:numerical_techniques}
The discussion to this point has outlined the formulation of the \methodAcronym\ method. We now consider numerical solution techniques. 
As \methodAcronym\ displays commonalities with optimal control problems, a large existing body of work can be leveraged. This section discuses a variety of such techniques and outlines how they can be used to solve the \methodAcronym\ system. We focus primarily on direct and indirect methods for \spatialAcronym\ trial subspaces, 
and then briefly outline direct and indirect methods for \spaceTimeAcronym\ trial subspaces. 
%Equation~\ref{eq:obj_gen_slab} displays commonalities with optimal control problems. For timeslabs $n = 1,\ldots,\nslabs$, the objective is to find the \textit{function} $\genstate^n(t)$ that minimizes the \textit{functional} $ \mathcal{J}^n$. The distinction between Eq.~\ref{eq:obj_gen_slab} and traditional optimal control problems is that, in the form presented in Eq.~\ref{eq:obj_gen_slab}, (1) there is not a controller and (2) there is no system that describes the dynamics for the generalized coordinates. 
%Techniques used to solve optimal control problems can be leveraged to solve Eq.~\ref{eq:obj_gen_slab} (or, equivalently, Eqs.~\eqref{eq:tclspg_oc1} and~\eqref{eq:tclspg_oc2}).

\subsection{\spatialAcronym: Direct and Indirect Methods}
Numerical techniques to
solve optimal control problems can be classified as either
\textit{direct} or \textit{indirect}
methods~\cite{conway_optimalcontrolreview}. Direct method eschew the calculus of variations 
and attempt to ``directly" solve the optimization problem~\eqref{eq:tclsrm} by first
numerically discretizing the system and objective functional and ``transcribing"
the infinite dimensional problem into a finite-dimensional nonlinear
programming (optimization) problem. Direct approaches thus ``discretize then optimize."
Indirect methods, on the other hand, leverage the calculus of variations to
derive the Euler--Lagrange equations, which comprise the first-order optimality conditions~\eqref{eq:lspg_continuous}-\eqref{eq:lspg_bcs} associated with the objective 
functional. Indirect methods then solve (numerically) the Euler--Lagrange equations. Thus, indirect methods ``optimize then discretize," and 
solve the optimization problem
``indirectly". For both direct and indirect methods, a variety of
discretization/solution approaches are then possible. Collocation methods,
finite elements, spectral methods, shooting methods, etc., are all examples of
plausible solution techniques.

>>>>

\section{Numerical solution techniques}\label{sec:numerical_techniques}
The discussion to this point has outlined the formulation of the \methodAcronym\ method. We now consider numerical solution techniques. 
As \methodAcronym\ displays commonalities with optimal control problems, a large existing body of work can be leveraged. This section discuses a variety of such techniques and outlines how they can be used to solve the \methodAcronym\ system. We focus primarily on direct and indirect methods for \spatialAcronym\ trial subspaces, 
and then briefly outline direct and indirect methods for \spaceTimeAcronym\ trial subspaces. 
%Equation~\ref{eq:obj_gen_slab} displays commonalities with optimal control problems. For timeslabs $n = 1,\ldots,\nslabs$, the objective is to find the \textit{function} $\genstate^n(t)$ that minimizes the \textit{functional} $ \mathcal{J}^n$. The distinction between Eq.~\ref{eq:obj_gen_slab} and traditional optimal control problems is that, in the form presented in Eq.~\ref{eq:obj_gen_slab}, (1) there is not a controller and (2) there is no system that describes the dynamics for the generalized coordinates. 
%Techniques used to solve optimal control problems can be leveraged to solve Eq.~\ref{eq:obj_gen_slab} (or, equivalently, Eqs.~\eqref{eq:tclspg_oc1} and~\eqref{eq:tclspg_oc2}).

\subsection{\spatialAcronym: direct and indirect methods}
Numerical techniques to
solve optimal control problems can be classified as either
\textit{direct} or \textit{indirect}
methods~\cite{conway_optimalcontrolreview}. Direct method eschew the calculus of variations 
and attempt to ``directly" solve the optimization problem~\eqref{eq:tclsrm} by first
numerically discretizing the system and objective functional and ``transcribing"
the infinite dimensional problem into a finite-dimensional nonlinear
programming (optimization) problem. Direct approaches thus ``discretize then optimize."
Indirect methods, on the other hand, leverage the calculus of variations to
derive the Euler--Lagrange equations, which comprise the first-order optimality conditions~\eqref{eq:lspg_continuous}-\eqref{eq:lspg_bcs} associated with the objective 
functional. Indirect methods then solve (numerically) the Euler--Lagrange equations. Thus, indirect methods ``optimize then discretize," and 
solve the optimization problem
``indirectly". For both direct and indirect methods, a variety of
discretization/solution approaches are then possible. Collocation methods,
finite elements, spectral methods, shooting methods, etc., are all examples of
plausible solution techniques.

### 2019-10-02 section4/spacetime_numerical.tex
p 6

\subsection{\spaceTimeAcronym\ Trial Spaces: Direct and Indirect Methods}
We now briefly consider \spaceTimeAcronym\ trial subspaces. 
For \spaceTimeAcronym\ trial subspaces, the distinction between a direct and indirect method is less clear as the optimization variable (i.e., the generalized coordinates) are finite 
dimensional. There is no need to transcribe an infinite dimensional optimization variable into a finite dimensional one. There is, however, still a need to develop a finite 
dimensional representation of the objective function~\eqref{eq:obj_gen_slab}. In what follows, we describe two techniques to do this: one that works with the FOM O$\Delta$E and one that works with the FOM ODE. We associate direct methods as those 
that work with the O$\Delta$E and indirect methods with those that work directly with the continuous ODE.

\subsection{\spaceTimeAcronym\ Trial Spaces: Direct Solution Technique}
The direct solution technique seeks to minimize the fully discrete objective function associated with the O$\Delta$E. For brevity, we again focus on linear multistep methods.

@@

\subsection{\spaceTimeAcronym\ Trial Spaces: Indirect Solution Technique}

@@

\subsection{\spaceTimeAcronym\ Trial Space: Summary}
As was just shown, \spaceTimeAcronym\ trial subspaces yield a series of space--time algebraic systems over each window. As a variety of work has examined space--time reduced-order models with \spaceTimeAcronym\ trial subspaces, 
a detailed exposition of solution techniques for these systems is not pursued here. It is sufficient to say that the 
space--time trial subspace yields a series of dense (non)linear systems to be solved over each window.

>>>>

\subsection{\spaceTimeAcronym\ trial spaces: direct and indirect methods}
We now briefly consider \spaceTimeAcronym\ trial subspaces. 
For \spaceTimeAcronym\ trial subspaces, the distinction between a direct and indirect method is less clear as the optimization variable (i.e., the generalized coordinates) are finite 
dimensional. There is no need to transcribe an infinite dimensional optimization variable into a finite dimensional one. There is, however, still a need to develop a finite 
dimensional representation of the objective function~\eqref{eq:obj_gen_slab}. In what follows, we describe two techniques to do this: one that works with the FOM O$\Delta$E and one that works with the FOM ODE. We associate direct methods as those 
that work with the O$\Delta$E and indirect methods with those that work directly with the continuous ODE.

\subsection{\spaceTimeAcronym\ trial spaces: direct solution technique}
The direct solution technique seeks to minimize the fully discrete objective function associated with the O$\Delta$E. For brevity, we again focus on linear multistep methods.

@@

\subsection{\spaceTimeAcronym\ trial spaces: indirect solution technique}

@@

\subsection{\spaceTimeAcronym\ trial space: summary}
As was just shown, \spaceTimeAcronym\ trial subspaces yield a series of space--time algebraic systems over each window. As a variety of work has examined space--time reduced-order models with \spaceTimeAcronym\ trial subspaces, 
a detailed exposition of solution techniques for these systems is not pursued here. It is sufficient to say that the 
space--time trial subspace yields a series of dense (non)linear systems to be solved over each window.
